---
title: "Project 1"
author: "Bruno de Melo"
date: "2/23/2020"
output: html_document
---
## Chess Cross Table Example
Idea is to combine the relevant lines/row into one, transform the dataset into a structured table/frame and work from there to calculate the opponents' average.

## Step 1
Read file, do some manipulation to clean the data and save the clean dataset into a file, in order to easily convert it to a structured table. I don't delete the separator "|", as it will help later in converting the table into a structured data frame.

```{r}
# open connection and read lines
library(tidyverse)
con <- file('https://raw.githubusercontent.com/bsvmelo/CUNY/master/tournamentinfo.txt', open='r')
w<-readLines(con)

head(w,8)
leng<-length(w)

#for loop in the lines collection and do some cleaning

for (i in 1:leng)
{
  if( i%%3 == 0 )
     {
         a<-w[i-1]
         b<-w[i]

# cleaning 
        c<-paste(a,b)
        c<-str_replace_all(c,"/","|")
        c<-str_replace_all(c,"->","|")
        c<-str_replace_all(c,"/ R:","|")
        c<-str_replace_all(c,"R:","")
        c<-str_trim(c, side="both")

# writing the cleaning data set into a temp file
        write(c,file="tempFile.txt", append=TRUE)
  }
}        

# close connection
        close(con)

# assign structured data back to a table within R
        
        out<-read.table("tempFile.txt",header = TRUE, sep="|")
       
        head(out,5)
```        
## Step 2
Here I create 2 temporary tables containing the main data that will populate the final output and also data containing the Rounds data, indexed by the player's number.
Since both tables have the same number of rows, it will be easier to make a reference when calculating the average.

```{r}
# temporary tables

        table1<-out[,c(1,2,11,3,13)]
        table1$Average<-0
        table2<-out[,c(1,4:10)] 
# extracting non relevant data from the pre rating column
        
        table1$Rtg..Pre<-str_extract_all(table1$Rtg..Pre,"[[:digit:]]{3,4}","")
```
## Step 3
This is the main for loop that checks in the second table, whether the player had a win, loss or draw, and looks for the respective opponent rating, and calculate the average      

Logic is to go element by element, check for W,D,L, and look for the corresponding opponent number, and store his/her rating in a variable   

``` {r}       
        for (i in 1:nrow(table2))
        {
          nogames<-0
          oppRtg<-0
          result<-0
          for (j in 2:ncol(table2)) 
          {
            if(substr(table2[i,j],1,1) == "W" | substr(table2[i,j],1,1) == "L" | substr(table2[i,j],1,1) == "D")
            {
              nogames <- nogames +1
              # here I'm extracting the numeric value from the string "W  39" while taking advantage of the fixed length
              noempty<- str_count(table2[i,j]," ")
              opp<-substr(table2[i,j],noempty+2,5)
              #here I store the opponent rating
              oppRtg<-oppRtg+as.numeric(table1[opp,5])
            }
          }
          # here the average is calculated and stored in table 1
          result<-oppRtg/nogames
          table1[i,6]<-round(result,0)
        }
```        
## Step 4
Final formatting to have the table ready to be exported to csv
```{r}
        final_table<-table1[,c(2:6)] 
        names(final_table)<-c("Player’s Name","Player’s State","Total Number of Points","Player’s Pre-Rating","Average Pre Chess Rating of Opponents") 
        write.csv(final_table, file="Project1-Bruno de Melo.csv")
        head(final_table)
        
```


