---
title: "HW 9"
author: "Bruno de Melo"
date: "3/29/2020"
output: html_document
---

## New York Times API

I'm using jsonlite package to convert results in a flat list and then convert to a data frame.

Two APIs have been queries:
Book review
Article search

But first things first, installed/calling relevant libraries.

```{r}

library(jsonlite)
library(tidyverse)

nyt_key_book_review<-"6SzKSy3JLBQhJRA2sie1z0fZSiO9kYGJ"
nyt_key_article_search<-"oREdUFLfvSwvdkb5iQj4GqvJ6aiqpa5"


```

## Book review

The Books API provides information about book reviews and The New York Times Best Sellers lists.

I'm getting book reviews on one of my favorite authors: Iain Banks.

```{r}
# Reading JSON files from NYT API and converting to a data.frame

# Book review
reviews_df <- fromJSON("https://api.nytimes.com/svc/books/v3/reviews.json?author=Iain+Banks&api-key=6SzKSy3JLBQhJRA2sie1z0fZSiO9kYGJ",flatten = TRUE) %>% data.frame()

# Printing
head(reviews_df)

```

## Article searching

As another example, I'm also getting articles mentioning Mr Banks. This data frame would require more work, giving that it searches for all string.

```{r}

# Article search
article_df <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=iain+m.+banks&api-key=joREdUFLfvSwvdkb5iQj4GqvJ6aiqpa5", flatten = TRUE) %>% data.frame()

# Printing
head(article_df,10)

```
