---
title: "Week 7"
author: "Bruno de Melo"
date: "3/14/2020"
output: html_document
---

## Working with HTML, XML and JSON in R

#Location

These are the files used in this assignment.

<a href="https://raw.githubusercontent.com/bsvmelo/CUNY/master/Sci-Fi_bookshelf1.html">Link to HTML file</a>

<a href="https://raw.githubusercontent.com/bsvmelo/CUNY/master/Sci-Fi_bookshelf1.xml">Link to XML file</a>

<a href="https://raw.githubusercontent.com/bsvmelo/CUNY/master/sci-fi_bookshelf2.json">Link to JSON file</a>



#Loading libraries

```{r}
if (!require('XML')) install.packages('XML')
if (!require('jsonlite')) install.packages('jsonlite')
library(XML)
library(jsonlite)
```

#1:HTML

#Reading HTML file

```{r}
# Reading HTML file from github
con <-'https://raw.githubusercontent.com/bsvmelo/CUNY/master/Sci-Fi_bookshelf1.html'
bookshelf_raw<-readLines(con, warn = FALSE)
# Printing
head(bookshelf_raw)
```

#Parsing HTML

```{r}
parsed_bookshelf<-htmlParse(bookshelf_raw)
print(parsed_bookshelf)
```

#Converting to R data.frame

```{r}

df_bookshelf <- readHTMLTable(parsed_bookshelf, stringsAsFactors=FALSE)

df_bookshelf[[1]]
```


#Displaying

```{r}
names(df_bookshelf[[1]])

df_bookshelf[[1]]$Title
```


#2:XML

#Reading XML file

```{r}
# Reading XML file from github
con <-'https://raw.githubusercontent.com/bsvmelo/CUNY/master/Sci-Fi_bookshelf1.xml'
bookshelf_raw_xml<-readLines(con, warn = FALSE)
# Printing
head(bookshelf_raw_xml)
```

#Parsing XML

```{r}
parsed_bookshelf_xml<-xmlParse(bookshelf_raw_xml)
print(parsed_bookshelf_xml)
```

#Converting to R data.frame

```{r}

(df_bookshelf_xml <- xmlToDataFrame(parsed_bookshelf_xml))

```

#Displaying

```{r}
df_bookshelf_xml$Title

df_bookshelf_xml$`Publication-Date`

df_bookshelf_xml$Author

```


#3:JSON

#Reading JSON

```{r}
# Reading JSON file from github
con <-'https://raw.githubusercontent.com/bsvmelo/CUNY/master/sci-fi_bookshelf2.json'
bookshelf_raw_js<-readLines(con, warn = FALSE)
# Printing
head(bookshelf_raw_js)
```

#Converting to R data.frame

```{r}
bookshelf_js<-fromJSON(bookshelf_raw_js)
bookshelf_js
```

#Displaying

```{r}
bookshelf_js$`Sci-Fi_bookshelf`$Title

bookshelf_js$`Sci-Fi_bookshelf`$Publisher
```


## Conclusion

All data frames are identical.