---
title: "Week 2"
author: "Bruno de Melo"
date: "2/9/2020"
output: html_document
---


## SQL  and R
## Movies Rating
## Installing packages and libraries
I could not find a way to make 100% reproducible mainly due to username/password. My solution was to create a super-user in MySQL who does not have a password.

```{r}
#install.packages("RMySQL")
#install.packages("dplyr")
library(RMySQL)
library(DBI)
library(dplyr)

db_user <- 'newuser'
db_password <- NULL
db_name <- NULL 
db_host <- NULL 
db_port <- 3306

drv<-dbDriver("MySQL")
movies <-  dbConnect(drv, user = db_user, password = db_password, dbname = db_name, host = db_host, port = db_port)
```

# Create database
Database Movie_ratings is created

```{r}
dbSendQuery(movies, "CREATE DATABASE movie_ratings;")

dbSendQuery(movies, "USE movie_ratings")
```
# Create table
Table ratings is created
```{r}
query<-"CREATE TABLE ratings (Friends TEXT, Movie TEXT, Rating TEXT);"

#Send the query to MySQL for execution
results <- dbSendQuery(movies, query)
dbClearResult(results)
```

# Insert data into tables - Brute Force Method
```{r}
query <- "INSERT INTO 
    ratings(Friends, Movie, Rating)
VALUES
('Friend1','Avengers: Endgame','5'),
('Friend1','The Lion King','10'),
('Friend1','Star Wars: The Rise of Skywalker',NULL),
('Friend1','Joker',NULL),
('Friend1','Spider-Man: Far From Home',NULL),
('Friend1','Jumanji: The Next Level',NULL),
('Friend2','Avengers: Endgame',NULL),
('Friend2','The Lion King',NULL),
('Friend2','Star Wars: The Rise of Skywalker','8'),
('Friend2','Joker','8'),
('Friend2','Spider-Man: Far From Home','9'),
('Friend2','Jumanji: The Next Level',NULL),
('Friend3','Avengers: Endgame','10'),
('Friend3','The Lion King',NULL),
('Friend3','Star Wars: The Rise of Skywalker',NULL),
('Friend3','Joker',NULL),
('Friend3','Spider-Man: Far From Home',NULL),
('Friend3','Jumanji: The Next Level',NULL),
('Friend4','Avengers: Endgame','5'),
('Friend4','The Lion King','7'),
('Friend4','Star Wars: The Rise of Skywalker','9'),
('Friend4','Joker','9'),
('Friend4','Spider-Man: Far From Home','4'),
('Friend4','Jumanji: The Next Level','6'),
('Friend5','Avengers: Endgame',NULL),
('Friend5','The Lion King',NULL),
('Friend5','Star Wars: The Rise of Skywalker',NULL),
('Friend5','Joker',NULL),
('Friend5','Spider-Man: Far From Home','5'),
('Friend5','Jumanji: The Next Level','9');"
results <- dbSendQuery(movies, query)
dbClearResult(results)
```
## Fetching dataset
```{r}
rs = dbSendQuery(movies, "select * from ratings")
datar = fetch(rs, n=-1)
dbClearResult(rs)
```
## Displaying table in R
```{r}
head(datar)
```
## Dealing with Missing Values or NAs
You also can find the sum and the percentage of missings in your dataset with the code below:
```{r}
Number_of_missing_ratings<-sum(is.na(datar))
Pct_missing_ratings<-mean(is.na(datar))
Number_of_missing_ratings
Pct_missing_ratings

```
## This function remove the missings
```{r}
na.omit(datar)
```


