---
title: "Week 5"
author: "Bruno de Melo"
date: "2/28/2020"
output: html_document
---

# Assignment 5 Tidying and Transforming Data

Objective is to tidy and transform data in order to perform analysis to compare the arrival delays of two airlines.

File is displayed below.


```{r}

library(tidyverse)
library(scales)

# Reading csv file, using [-3,] parameter to delet blank line between rows with data
air1 <- read.csv(file = 'https://raw.githubusercontent.com/bsvmelo/CUNY/master/airline_data.csv')[-3,]

head(air1)

```
# Tidying and transforming data

Step 1: Changing column names.

Step 2: Filling blank cells with Airline name.
```{r}
#Step 1
names(air1)[1:2]<-c("Airline","Status")
#Step 2
air1[2,1]<-air1[1,1]
air1[4,1]<-air1[3,1]
head(air1)
```
Step 3: Using Gather() function to transform column with city names into a new variable. 
```{r}
#Step 3
air2<-gather(air1[,3:7],key="City", value="No_Flights")
head(air2)
```
Step 4: Transpose columns Airline and Status to newly created data frame - air2.

```{r}
#Step 4
air2[,3:4]<-air1[,1:2]
air2[,4]<-str_trim(air2[,4])

head(air2)

```

# Exploratory Data Analysis

For a quick visual inspection, plot below shows Number of flights by Airline by City. 


```{r}
ggplot(air2, aes(fill=City, x=Airline, y=No_Flights)) + geom_col(position="stack") + ylab("Number of Flights")
```

Some highlights:

1: AM West has more flights than Alaska.

2: AM West has more flights to Phoenix than to any other city.

3: Alaska has more flights to Seattle than to any other city.


Another visualization, showing Number of flights by Airline by Status
```{r}
ggplot(air2, aes(fill=Status, x=Airline, y=No_Flights)) + geom_col(position="stack") + ylab("Number of Flights")

```

Conclusion:
Visually, it seems that AM West might have more delayed arrivals than Alaska, in aggregate. However, we need to make each set of stacked bars the same height, for a proper comparison.

```{r}
ggplot(air2, aes(fill=Status, x=Airline, y=No_Flights)) + geom_col(position="fill") + ylab("Number of Flights")
```


My first impression is not correct! AM West is more punctual than Alaska.


# EDA - by City


I now show a multiple bar charts tiling showing Status breakdown by City by Airlines for a quick visual comparative inspection. 

```{r}
ggplot(air2, aes(fill=Status, x=Airline, y=No_Flights)) + geom_col() + ggtitle("Status by City") + facet_wrap(~City)

```

Some highlights:

1: AM West flies more frequently to Phoenix than any other location
2: Alaska flies more frequently to Seattle than any other location

As before, we need to make each set of stacked bars the same height, for a proper comparison.

```{r}
ggplot(air2, aes(fill=Status, x=Airline, y=No_Flights)) + geom_col(position="fill") + ggtitle("Status by City") + facet_wrap(~City)

```

Some highlights:

1: City with most delayed arrival is San Francisco, followed by Seattle, on both airlines as it seems visually.

2: As seem above, AM West biggest destination is Phoenix but this fact doesn't translate into more delays. 

3: On Alaska, biggest destination is Seattle and this seems to impact delayed frequency.

Calculation showing the percentage of delayed flights will done in the next step. 


# Summarizing data using piping

In this step, I will aggregate data in order to perform a calculation that will show which airline is more punctual.

```{r}
# Total fligths by status
summary_total <- air2 %>%
  group_by(Airline, Status) %>%
  summarize(Total_Flights=sum(No_Flights))%>%
  spread(key="Status", value="Total_Flights") %>%
  mutate(total=`on time`+delayed,`On_Time`=percent(round((`on time`/total),2)),`Delayed`=percent(round((delayed/total),2))) %>%
  arrange(desc(`On_Time`))
head(summary_total)
```
# Conclusion 1

1: As confirmation to the visual inspection above, AM West is indeed more punctual than Alaska by 2 percentage points!

Next analysis, will look into % Delayed by city by airline, in aggregate.
```{r}
# best by City
# air2<-air2[,c(2,1,3,4)]
summary_city <- air2 %>%
  group_by(City,Airline,Status) %>%
  summarize(Total_Flights=sum(No_Flights))%>%
  spread(key="Status", value="Total_Flights") %>%
  mutate(total=`on time`+delayed,`% On_Time`=percent(round((`on time`/total),2)),`% Delayed`=percent(round((delayed/total),2))) %>%
  arrange(`% On_Time`, City)
# Top 10 delayed cities
head(summary_city, n=10)
```
# Conclusion 2

2: Top 3 cities with more delayed flights are San Francisco and Seattle on both AM West and Alaska, as suspected from the charts above.

On Alaska, second city with more delayed flights is also Seattle.


Looking into a similar table but grouping by Airline

```{r}
summary_airline_1 <- air2 %>%
  group_by(Airline,City, Status) %>%
  summarize(Total_Flights=sum(No_Flights))%>%
  spread(key="Status", value="Total_Flights") %>%
  mutate(total=`on time`+delayed,`% On_Time`=percent(round((`on time`/total),2)),`% Delayed`=percent(round((delayed/total),2))) %>%
  arrange(Airline,`% On_Time`, City)
# Top 10 delayed cities
head(summary_airline_1, n=10)
```
# Conclusion 3

3: As seem before, Alaska flies more to Seattle than to any other location and has the second worst record in terms of delayed flights.
On AM West, despite having more fligths to Phoenix than any other location, this is the most punctual destination.


