---
title: "HW 9"
author: "Bruno de Melo"
date: "3/29/2020"
output: html_document
---

## New York Times API

I'm using jsonlite package to convert results in a flat list and then convert to a data frame.

Two APIs have been queries:
Book review
Article search

But first things first, installed/calling relevant libraries.

```{r}
library(XML)
library(jsonlite)
library(tidyverse)

nyt_key_book_review<-"6SzKSy3JLBQhJRA2sie1z0fZSiO9kYGJ"
nyt_key_article_search<-"oREdUFLfvSwvdkb5iQj4GqvJ6aiqpa5"


```

## Book review

The Books API provides information about book reviews and The New York Times Best Sellers lists.

I'm getting book reviews on one of my favorite authors: Iain Banks.

```{r}
# Reading JSON files from NYT API and converting to a data.frame

# Book review
reviews_df <- fromJSON("https://api.nytimes.com/svc/books/v3/reviews.json?author=Iain+Banks&api-key=6SzKSy3JLBQhJRA2sie1z0fZSiO9kYGJ",flatten = TRUE) %>% data.frame()

# Printing
head(reviews_df)



```

# Goodreads API

I would like to check whether any of these books reviewed by the Times are part of the famous Culture series.

To do so, I will call an API from Goodreads.com: https://www.goodreads.com/api.

There are two API calls:
First: Author name to grab author id
Second: Series list, that uses author id

Steps:
1: Finding author id

```{r}
# connection
# author name to be parsed in the URL
con <-'https://www.goodreads.com/api/author_url/Iain%20M%20Banks?key=UZDsXRdqqW0GHin6a2Iw'
author_xml<-readLines(con, warn = FALSE)
# XML parsing
author_xml<-xmlParse(author_xml)
# grabbing author id
author_id <- xpathSApply(author_xml, "//*/author/@id")
y<-str_trim(author_id, side = c("both"))
```

This is Mr Banks' id: `r y`.

2: Extract author's series
```{r}
# connection
# author id to be parsed in the URL
url_gr<-paste("https://www.goodreads.com/series/list/",y,".xml?key=UZDsXRdqqW0GHin6a2Iw",sep="")

con1<-url_gr
# XML parsing
series_xml<-readLines(con1, warn = FALSE)
head(series_xml,10)
series_xml<-xmlParse(series_xml)

# extracting Book titles into a data frame
title_df <- xmlToDataFrame(nodes = getNodeSet(series_xml, "//*/best_book"), stringsAsFactors = FALSE)

# Displaying Book titles
head(title_df$title,20)
```

3: Comparing NYT book list to Goodreads list.

```{r}

result<-str_detect(title_df$title,reviews_df$results.book_title)

```

Unfortunately, none of the books reviewed by NYT belong to the Culture series.

`r result`

## Article searching

As another example, I'm also getting articles mentioning Mr Banks. This data frame would require more work, giving that it searches for all string.

```{r}

# Article search
article_df <- fromJSON("http://api.nytimes.com/svc/search/v2/articlesearch.json?q=iain+m.+banks&api-key=joREdUFLfvSwvdkb5iQj4GqvJ6aiqpa5", flatten = TRUE) %>% data.frame()

# Printing
head(article_df,10)

```
